# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r,echo=TRUE}
<<<<<<< HEAD
unzip("activity.zip",list=T)
=======
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868
unzip("activity.zip")
d <- read.csv("activity.csv",header=T)
```
## What is mean total number of steps taken per day?

```{r,echo=TRUE}
<<<<<<< HEAD
steps_taken_per_day <- tapply(d$steps,as.factor(d$date),sum)
hist(steps_taken_per_day)
mean(steps_taken_per_day,na.rm = T)
median(steps_taken_per_day,na.rm = T)
```

Mean: 10766.19
Median: 10765
=======
hist(tapply(d$steps,as.factor(d$date),mean))
mean(tapply(d$steps,as.factor(d$date),mean),na.rm = T)
median(tapply(d$steps,as.factor(d$date),mean),na.rm = T)
```

The mean and median are 37.38
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868

## What is the average daily activity pattern?

```{r,echo=TRUE}
require("ggplot2")
require("plyr")
summary_by_interval <- ddply(d,.(interval),summarize,Total=sum(steps,na.rm=T),Mean=mean(steps,na.rm=T),Median=median(steps, na.rm=T))
<<<<<<< HEAD
qplot(data = summary_by_interval,x=interval, y=Mean,geom = "line", ylab = "Average number of steps over the 2 months")
# Looking up the 5-minute interval that, on average, contains the maximum number of steps
=======
ggplot(summary_by_interval, aes(x=interval, y=Mean)) + geom_line()
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868
summary_by_interval$interval[summary_by_interval$Mean == max(summary_by_interval$Mean)]
```
The interval 835 has the highest average number of steps across all days for the data.

## Imputing missing values

<<<<<<< HEAD
1. There are 2304 NA values in original dataset
```{r,echo=TRUE}
summary(d)
```
2. To 'fill in' the NA values in this dataset, I will use the mean value for the given interval. This will be done by using a for loop to look at each row. For each row, if the steps value is NA, the function will look up the mean value for that interval in a summary table that has the total number or steps, mean, and median calculated for each interval.
=======
1. There are 2304 NA values in out dataset
```{r,echo=TRUE}
summary(d)
```
2. To 'fill in' the NA values in this dataset, I will use the median value for the given interval. 
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868

3. New dataset with the NA's filled in using the mean for the interval
```{r,echo=TRUE}
summary_by_interval <- ddply(d,.(interval),summarize,Total=sum(steps,na.rm=T),Mean=mean(steps,na.rm=T),Median=median(steps, na.rm=T))
<<<<<<< HEAD
# d2 will be my new table with all NA values replaced with mean values.
=======
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868
d2 <- d
for (i in 1:nrow(d2)) {
  if(is.na(d2[i,1])) {
    d2[i,1] <- summary_by_interval$Mean[summary_by_interval$interval == d2[i,3]]
  }
}
head(d2)
summary(d2)
```

<<<<<<< HEAD
4. The difference between the mean and median for the data set with the missing values and the data set without is very small.

```{r,echo=TRUE}
# The total number of steps per day using dataset without missing values.
steps_taken_per_day_filled_na <- tapply(d2$steps,as.factor(d2$date),sum)
hist(steps_taken_per_day_filled_na)
mean(steps_taken_per_day_filled_na)
median(steps_taken_per_day_filled_na)
# Calculating the difference between the mean and median for the dataset with missing values and the dataset without.
mean(steps_taken_per_day_filled_na) - mean(steps_taken_per_day,na.rm=T) 
median(steps_taken_per_day_filled_na) - median(steps_taken_per_day,na.rm=T) 
=======
4. The difference between the mean and median for the data set with the missing values and the date set is very close to zero.

```{r,echo=TRUE}
hist(tapply(d2$steps,as.factor(d2$date),mean))
mean(tapply(d2$steps,as.factor(d2$date),mean))
median(tapply(d2$steps,as.factor(d2$date),mean))
mean(tapply(d2$steps,as.factor(d2$date),mean)) - mean(tapply(d$steps,as.factor(d$date),mean),na.rm=T) 
median(tapply(d2$steps,as.factor(d2$date),mean)) - median(tapply(d$steps,as.factor(d$date),mean),na.rm=T) 
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868
```

## Are there differences in activity patterns between weekdays and weekends?

Overall, the trend seems to be that this individual moves more durring the weekend. However, looking at movement per interval, this individual moves the most in the morning on weekdays; we might assume this is their walk to work.

```{r,echo=TRUE}
require("timeDate")
<<<<<<< HEAD
# d2 is same dataframe as RAW data but with all NA's replaced with mean of interval (See Imputing missing values #3)
=======
>>>>>>> 764e19b0ef36239d278734bdc8023fa8a0929868
d2$typeday <- as.factor(isWeekday(d2$date))
levels(d2$typeday) <- list(weekend = "FALSE", weekday = "TRUE")
qplot(x = interval,y = steps, data = d2 , facets = (typeday~.),geom = "line",stat = 'summary', fun.y = 'mean', main = "Activity Patterns between weekdays vs weekends",xlab = "Interval (in 5min increments)", ylab = "Number of Steps (Averaged over 2 month period)")
```

